A Theory for Data Migration of Information Systems

### 2022-09-17 2022 Migration/articleMigration.tex
minor

	Now let us define an information system.
	To save writing in the sequel, we will write $a\ r\ b$ to denote that $\triple{a}{r}{b}\in\triples$.
	We define a constraint that states that a student can enroll for any module that is part of a course that student takes.
	In logic, this constraint is written as:
\[\begin{array}{l}\forall s\in\text{\tt Student}, m\in\text{\tt Module}\ \exists c\in\text{\tt Course}:\\
s\ \text{\tt isEnrolledFor}\ m\ \rightarrow\ s\ \text{\tt takes}\ c\ \wedge\ m\ \text{\tt isPartOf}\ c
\end{array}\]
	In Ampersand, which is a syntactically sugared form of relation algebra,
	we give that constraint a name and state a role to maintains it:
\begin{verbatim}
RULE EnrollRule: isEnrolledFor |- takes;isPartOf~
ROLE Administrator MAINTAINS EnrollRule
\end{verbatim}
	The signature of this rule is $\pair{\text{\tt Student}}{\text{\tt Module}}$, which is derived by the Ampersand compiler.
	Ampersand also derives the violation set as:
\[\begin{array}{l}
	\viol{\triples}{\text{\tt EnrollRule}}\\
	\hspace{1cm}=\{\pair{s}{m}|s\ \text{\tt isEnrolledFor}\ m\ \wedge\not\exists c: s\ \text{\tt takes}\ c\ \wedge\ m\ \text{\tt isPartOf}\ c\}
\end{array}\]
	So, next to a dataset we have one rule, one role, a relation $\maintain$ that contains one pair.
	So we have now defined an example information system.
\subsection{Data migration}
	To migrate information system $\infsys$ to $\infsys'$, we take the disjoint union $\infsys\sqcup\infsys'$ of both information systems
	and add rules to describe the migration.
	$\infsys$ is the existing system, which contains data to be preserved.
	$\infsys'$ is the new system, which contains no or hardly any data.
	It does, however, contain the concepts, relations, rules and roles of the new system.
\begin{figure}[bht]
	\begin{center}
	  \includegraphics[scale=.45]{Migration.png}
	\end{center}
\caption{Data migration}
\label{fig:event flow}
\end{figure}

>>>>

	Now let us define an information system.
	To save writing in the sequel, we will write $a\ r\ b$ to denote that $\triple{a}{r}{b}\in\triples$.
	We define a constraint that states that a student can enroll for any module that is part of a course that student takes.
	In logic, this constraint is written as:
\[\begin{array}{l}\forall s\in\text{\tt Student}, m\in\text{\tt Module}\ \exists c\in\text{\tt Course}:\\
s\ \text{\tt isEnrolledFor}\ m\ \rightarrow\ s\ \text{\tt takes}\ c\ \wedge\ m\ \text{\tt isPartOf}\ c
\end{array}\]
	In Ampersand, which is a syntactically sugared form of relation algebra,
	we give that constraint a name and state a role to maintains it:
\begin{verbatim}
RULE EnrollRule: isEnrolledFor |- takes;isPartOf~
ROLE Administrator MAINTAINS EnrollRule
\end{verbatim}
	The signature of this rule is $\pair{\text{\tt Student}}{\text{\tt Module}}$, which is derived by the Ampersand compiler.
	Ampersand also derives the violation set as:
\[\begin{array}{l}
	\viol{\triples}{\text{\tt EnrollRule}}\\
	\hspace{1cm}=\{\pair{s}{m}|s\ \text{\tt isEnrolledFor}\ m\ \wedge\not\exists c: s\ \text{\tt takes}\ c\ \wedge\ m\ \text{\tt isPartOf}\ c\}
\end{array}\]
	Notice that the current population yields an empty set of violations,
	so the rule is satisfied.
	Having one rule, one role, a relation $\maintain$ that contains one pair, and a dataset,
	we have now defined a complete information system (albeit small).

\subsection{Data migration}
	To migrate information system $\infsys$ to $\infsys'$, we take the disjoint union $\infsys\sqcup\infsys'$ of both information systems
	and add rules to describe the migration.
	$\infsys$ is the existing system, which contains data to be preserved.
	$\infsys'$ is the new system, which contains no or hardly any data.
	It does, however, contain the concepts, relations, rules and roles of the new system.
\begin{figure}[bht]
	\begin{center}
	  \includegraphics[scale=.45]{Migration.png}
	\end{center}
\caption{Data migration}
\label{fig:event flow}
\end{figure}
